The \dealii Library, Version 9.6

### 2024-09-19 9.6/paper.tex
Improve the orientation paragraphs.

    Since very early in the history of the library, the
    \texttt{AffineConstraints} class interface required building such
    constraints in multiple steps: First, one declared a degree of
    freedom as constrained; then one added the dependencies one after
    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
    14)$ in the hanging node example above); then one added
    inhomogeneities (by setting it to $42$ in the Dirichlet example
    above). This piecemeal approach is cumbersome and prevents the
    library from performing certain error checking steps because a
    constraint is not known to be completely built at any given point. The new
    \texttt{add\_constraint()} function now allows defining a
    constraint in one step.
  \item \dealii{} internally needs to keep track of the relative
    orientations of the coordinate systems associated with neighboring
    cells, as well as of the coordinate systems of faces and edges
    within the coordinate system of the cells they are part
    of. Historically, different places within the library that had to
    do that had grown their own, independent ways of representating
    this information, much of which was difficult to understand.
    We have now cleaned up this area. Specifically, instead of using three
    Booleans to define relative orientations for hexedra (\textit{orientation},
    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
     \texttt{char} -- i.e., the smallest integer data type that has at
     least three bits to represent the information. Because the way to represent
     orientations is part of the public interface, the change causes
     some incompatibilities in types and function signatures.
  \todo[inline]{DavidW: Please describe some details here. -- When
    done, please remove this comment.}
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.

>>>>

    Since very early in the history of the library, the
    \texttt{AffineConstraints} class interface required building such
    constraints in multiple steps: First, one declared a degree of
    freedom as constrained; then one added the dependencies one after
    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
    14)$ in the hanging node example above); then one added
    inhomogeneities (by setting it to $42$ in the Dirichlet example
    above). This piecemeal approach is cumbersome and prevents the
    library from performing certain error checking steps because a
    constraint is not known to be completely built at any given point. The new
    \texttt{add\_constraint()} function now allows defining a
    constraint in one step.
  \item We refactored the systems for managing relative line and face
    orientations to significantly improve consistency across several library
    modules, including finite element classes, periodicity, and the
    \texttt{p4est} interface. \dealii{} has supported using unstructured
    three-dimensional meshes for a long time. Features like discontinuous
    Galerkin methods and higher-order elements require that lines and faces of
    adjacent elements are consistently oriented. For example, consider a line
    containing two degrees of freedom, which, in reference coordinates, are
    located at $x_1 = \frac 1 3$ and $x_2 = \frac 2 3$. If the line is read from
    left-to-right then we get points $(x_1, x_2)$ whereas if it is read from
    right-to-left we get $(x_2, x_1)$. \dealii{} guarantees that this order will
    be consistent across all cells sharing that line by storing a flag
    indicating whether (relative to the vertex enumeration on the present cell)
    that particular line is in the standard or reversed orientation. Faces in 3d
    are treated in a similar way. The only significant difference between lines
    and 3d faces is that, instead of two possible relative orientations,
    triangular faces have six and quadrilateral faces have eight.

    Historically, different places within the library implemented their own
    ad-hoc orientation logic, data encodings, and default values. We cleaned up
    this subsystem and instead of using three different booleans
    (\emph{orientation}, \emph{rotation}, and \emph{flip}), bitsets, or several
    other encodings, we use a single \texttt{unsigned char} which we named the
    \emph{combined orientation}. Since the representation in some library
    interfaces has changed, most of the old interfaces have been deprecated and
    a few (such as the low-level interface to periodic boundary conditions) had
    to be incompatibly altered or removed.
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.
